Nested sampling with two objective functions

### 2015-12-04 Paper/ms.tex
A few more words

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, the method should satisfy
the following properties:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which the enclosed prior mass shrinks by a (roughly) constant and
known factor each iteration.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm, (as well as alternatives such as
Diffusive Nested Sampling \citep{dnest}) has these properties.

@@

Consider running standard NS with $N$ particles
using the objective function $L_1(\xx)$.
After $M$ iterations, we'd be able to approximate
the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
from $X\approx e^{-M/N}$ to $X=1$. Defining $Y=-\ln(X)$, we could also
construct an approximation to the function
$L(Y)$, a non-decreasing function of $Y$, from $Y=0$ to $Y\approx M/N$.

@@

The proposed algorithm is given in Algorithm~\ref{alg:algorithm}
and took three years of
experimentation to develop. Please contact the first author for a list of
failed ideas, some of which might lead to potential alternatives to this
algorithm. The basic idea is to try to mimic a standard NS run, where the
objective function being increased is
\begin{eqnarray}
Y(\xx) &=& -\ln\left[X\left(L_1(\xx), L_2(\xx)\right)\right].
\end{eqnarray}

>>>>

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be a variant of Nested Sampling, the algorithm should satisfy
the following properties:
\begin{enumerate}
\item It should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item It should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior $\pi(L_1, L_2)$
would typically imply.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which the enclosed prior mass shrinks by a (roughly) constant and
known factor each iteration.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm, (as well as alternatives such as
Diffusive Nested Sampling \citep{dnest}) has these properties but for a
single objective function.

@@

Consider the following proposal. Imagine running standard NS with $N$ particles
using the objective function $L_1(\xx)$.
After $M_1$ iterations, we'd be able to approximate
the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
from $X\approx e^{-M_1/N}$ to $X=1$. Defining $Y=-\ln(X)$, we could also
construct an approximation to the function
$L_1(Y)$, a non-decreasing function of $Y$, from $Y=0$ to $Y\approx M_1/N$.
After $M$ iterations, we keep the last value of $L_1(\xx)$ obtained
(call it $L_1^*$), and continue to impose the constraint $L_1(\xx) > L_1^*$.

Subject to the constraint $L_1(\xx) > L_1^*$, standard NS can continue but
attempting to increase $L_2$ instead of $L_1$. After $M_2$ such iterations,
the algorithm would have compressed the prior by a factor
$e^{(M_1 + M_2)/N}$ and therefore ended at a point in the $(L_1, L_2)$ plane
where $X(L_1, L_2) \approx e^{-(M_1 + M_2)/N}$. Any such run consisting of
$M_1 + M_2$ iterations will therefore appoximately finish on a contour line
of the function $X(L_1, L_2)$.

@@

The basic idea of our algorithm is to try to mimic a standard NS run, where the
objective function being increased is
\begin{eqnarray}
Y(\xx) &=& -\ln\left[X\left(L_1(\xx), L_2(\xx)\right)\right].
\end{eqnarray}
Standard NS cannot be applied simply because we cannot evaluate $X(L_1, L_2)$ and
therefore cannot restrict the parameter space based on the value of $X$ (or
equivalently $Y$. However, given a number of Monte Carlo samples, we
can approximate this.
